paper_replication

### 2021-12-15 paper_replication.Rmd
Update paper_replication.Rmd

```{r}
#run models
countydemanalysis <- countydem %>% filter(inanalysis == 1)

@@

```{r}
natality <- read.delim("natalitydata.txt")
maternalmortality <- read.delim("maternalmortality.txt")
infantdeath <- read.delim("infantdeathrecords.txt")

@@

sum(is.na(countydem$infantdeathrate)/nrow(countydem))
countydem$maternaldeathrate <- na_if(countydem$maternaldeathrate, "Suppressed")
sum(is.na(countydem$maternaldeathrate)/nrow(countydem))
sum(is.na(countydem$Birth.Rate)/nrow(countydem))

>>>>

```{r}
#run models
countydemanalysis <- df %>% filter(inanalysis == 1)

@@

```{r}
####### add in natality, maternalmortality, and infantdeath data #######
natality <- read.delim("natalitydata.txt")
maternalmortality <- read.delim("maternalmortality.txt")
infantdeath <- read.delim("infantdeathrecords.txt")

@@

sum(is.na(countydem$infantdeathrate)/nrow(countydem))
countydem$maternaldeathrate <- na_if(countydem$maternaldeathrate, "Suppressed")
countydem$maternaldeathrate <- na_if(countydem$maternaldeathrate, "Missing")

sum(is.na(countydem$maternaldeathrate)/nrow(countydem))
sum(is.na(countydem$Birth.Rate)/nrow(countydem))

```{r}
#### recode restriction level and include all counties with abortion data in analysis ####

countydem <- countydem %>% mutate(hasabortiondata = case_when(is.na(AbortionRate2010) ~ 0, 
                                                              !(is.na(AbortionRate2010)) ~ 1))

countydemanalysis <- countydem %>% filter(hasabortiondata == 1)

summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable
summarytable %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Abortion Data Included in Analysis by State") %>% kable_styling()
names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

countydem$hasabortiondatalabel <- factor(countydem$hasabortiondata, levels = c(1, 0), labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data"))

table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasabortiondatalabel, 
       data = countydem,
       extra.col = list(`P-value` = p_value_t_test),
       footnote = 'P-values are derived by two sample t-tests.'
       )

#calculate propensity scores
#include education this time
propensitywts <- glm(hasabortiondata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
summary(propensitywts)

ggplot(df, aes(x = ps, group = hasabortiondatalabel)) + geom_density(aes(fill = hasabortiondatalabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")

ggplot(df, aes(x = ps, fill = hasabortiondatalabel, color = hasabortiondatalabel)) +
  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
  facet_grid(as.factor(hasabortiondata) ~ .) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  scale_color_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical')

df %>%
  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
  group_by(hasabortiondata, ps.grp) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(n2 = ifelse(hasabortiondata == 0, yes = n, no =  -1*n)) %>%
  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasabortiondata))) +
  geom_bar(stat = 'identity', position = 'identity') +
  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasabortiondata == 0, 8, -8))) +
  xlab('Probability of Being in the Analysis') +
  ylab('N') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete(name = "", labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data")) +
  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
  theme(legend.position = 'bottom', legend.direction = 'vertical',
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())

#run model
countydemanalysis <- df %>% filter(hasabortiondata == 1)

model <- lm(AbortionRate2010 ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemanalysis, weights = ps)
summary(model)

```{r}
###further clean up maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "[()]", replacement = "")
countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "Unreliable", replacement = "")
countydem$maternaldeathrate <- as.numeric(countydem$maternaldeathrate)
hist(countydem$maternaldeathrate)

countydem$infantdeathrate <- na_if(countydem$infantdeathrate, "Missing County")
countydem$infantdeathrate <- gsub(countydem$infantdeathrate, pattern = "[()]", replacement = "")
countydem$infantdeathrate <- gsub(countydem$infantdeathrate, pattern = "Unreliable", replacement = "")
countydem$infantdeathrate <- as.numeric(countydem$infantdeathrate)
hist(countydem$infantdeathrate)

countydem$Birth.Rate <- na_if(countydem$Birth.Rate, "Missing County")
countydem$Birth.Rate <- as.numeric(countydem$Birth.Rate)
hist(countydem$Birth.Rate)

countydem$Fertility.Rate <- na_if(countydem$Fertility.Rate, "Missing County")
countydem$Fertility.Rate <- as.numeric(countydem$Fertility.Rate)
hist(countydem$Fertility.Rate)

```
```{r}
### repeat analysis for infant death rate ###
countydem <- countydem %>% mutate(hasinfantdeathdata = case_when(is.na(countydem$infantdeathrate) ~ 0, 
                                                              !(is.na(countydem$infantdeathrate)) ~ 1))

countydeminfantdeathanalysis <- countydem %>% filter(hasinfantdeathdata == 1)

summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(infantdeathrate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable %>% mutate(percent = infantdeathrate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Data Included in Analysis by State") %>% kable_styling()
names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

countydem$hasinfantdeathdatalabel <- factor(countydem$hasinfantdeathdata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))

p_value_t_test <- function(x, ...){
    y <- unlist(x[1:2])
    g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
    p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
    p_value <- ifelse(p_value == '0', '< 0.001', p_value)
}

print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasinfantdeathdatalabel, 
       data = countydem,
       extra.col = list(`P-value` = p_value_t_test),
       footnote = 'P-values are derived by two sample t-tests.'
       ))

#calculate propensity scores
#include education this time
propensitywts <- glm(hasinfantdeathdata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())

plot1 <- ggplot(df, aes(x = ps, group = hasinfantdeathdatalabel)) + geom_density(aes(fill = hasinfantdeathdatalabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")
print(plot1)

plot2 <- ggplot(df, aes(x = ps, fill = hasinfantdeathdatalabel, color = hasinfantdeathdatalabel)) +
  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
  facet_grid(as.factor(hasinfantdeathdata) ~ .) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  scale_color_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical')
print(plot2)

plot3 <- df %>%
  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
  group_by(hasinfantdeathdata, ps.grp) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(n2 = ifelse(hasinfantdeathdata == 0, yes = n, no =  -1*n)) %>%
  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasinfantdeathdata))) +
  geom_bar(stat = 'identity', position = 'identity') +
  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasinfantdeathdata == 0, 8, -8))) +
  xlab('Probability of Being in the Analysis') +
  ylab('N') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete(name = "", labels = c("Counties in States With Data", "Counties in States Without Data")) +
  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
  theme(legend.position = 'bottom', legend.direction = 'vertical',
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
print(plot3)

#run model
countydeminfantdeathanalysis <- df %>% filter(hasinfantdeathdata == 1)

model <- lm(infantdeathrate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydeminfantdeathanalysis, weights = ps)
summary(model)

```{r}
### repeat analysis for birth rate ###

countydem <- countydem %>% mutate(hasbirthratedata = case_when(is.na(countydem$Birth.Rate) ~ 0, 
                                                              !(is.na(countydem$Birth.Rate)) ~ 1))

countydembirthrateanalysis <- countydem %>% filter(hasbirthratedata == 1)

summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(Birth.Rate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
usedcounties
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable
summarytable %>% mutate(percent = Birth.Rate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Data Included in Analysis by State") %>% kable_styling()
names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

countydem$hasbirthratedatalabel <- factor(countydem$hasbirthratedata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))

p_value_t_test <- function(x, ...){
    y <- unlist(x[1:2])
    g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
    p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
    p_value <- ifelse(p_value == '0', '< 0.001', p_value)
}

print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasbirthratedatalabel, 
       data = countydem,
       extra.col = list(`P-value` = p_value_t_test),
       footnote = 'P-values are derived by two sample t-tests.'
       ))

#calculate propensity scores
#include education this time
propensitywts <- glm(hasbirthratedata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
summary(propensitywts)

p.score <- predict(propensitywts, countydem, type = "response")
df <- countydem
df$ps <- p.score
#3 plots of distribution of propensity scores

plot1 <- ggplot(df, aes(x = ps, group = hasbirthratedatalabel)) + geom_density(aes(fill = hasbirthratedatalabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")
print(plot1)

plot2 <- ggplot(df, aes(x = ps, fill = hasbirthratedatalabel, color = hasbirthratedatalabel)) +
  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
  facet_grid(as.factor(hasbirthratedata) ~ .) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  scale_color_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical')
print(plot2)

plot3 <- df %>%
  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
  group_by(hasbirthratedata, ps.grp) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(n2 = ifelse(hasbirthratedata == 0, yes = n, no =  -1*n)) %>%
  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasbirthratedata))) +
  geom_bar(stat = 'identity', position = 'identity') +
  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasbirthratedata == 0, 8, -8))) +
  xlab('Probability of Being in the Analysis') +
  ylab('N') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete(name = "", labels = c("Counties in States With Data", "Counties in States Without Data")) +
  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
  theme(legend.position = 'bottom', legend.direction = 'vertical',
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
print(plot3)

#run model
countydembirthrateanalysis <- df %>% filter(hasbirthratedata == 1)

model <- lm(Birth.Rate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydembirthrateanalysis, weights = ps)
summary(model)

test <- countydembirthrateanalysis %>% filter(ps > 0.9999999)
```

```{r}
### repeat analysis for fertility rate ###

countydem <- countydem %>% mutate(hasfertilityratedata = case_when(is.na(countydem$Fertility.Rate) ~ 0, 
                                                              !(is.na(countydem$Fertility.Rate)) ~ 1))

countydemfertilityrateanalysis <- countydem %>% filter(hasfertilityratedata == 1)

summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(Fertility.Rate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
usedcounties
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable
summarytable %>% mutate(percent = Fertility.Rate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Data Included in Analysis by State") %>% kable_styling()
names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

countydem$hasfertilityratedatalabel <- factor(countydem$hasfertilityratedata, levels = c(1, 0), labels = c("Counties in States With Data", "Counties in States Without Data"))

p_value_t_test <- function(x, ...){
    y <- unlist(x[1:2])
    g <- factor(rep(1:length(x[1:2]), times=sapply(x[1:2], length)))
    p_value <- t.test(y ~ g)$p.value %>% round(3) %>% as.character()
    p_value <- ifelse(p_value == '0', '< 0.001', p_value)
}

print(table1(~ Women + White.pct + Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + College.Grad.pct + percentdem2008 | hasfertilityratedatalabel, 
       data = countydem,
       extra.col = list(`P-value` = p_value_t_test),
       footnote = 'P-values are derived by two sample t-tests.'
       ))

#calculate propensity scores
#include education this time
propensitywts <- glm(hasfertilityratedata ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008 + College.Grad.pct, data=countydem, family=binomial())
summary(propensitywts)

p.score <- predict(propensitywts, countydem, type = "response")
df <- countydem
df$ps <- p.score
#3 plots of distribution of propensity scores

plot1 <- ggplot(df, aes(x = ps, group = hasfertilityratedatalabel)) + geom_density(aes(fill = hasfertilityratedatalabel), alpha = 0.2) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical') +
  ylab("Density")
print(plot1)

plot2 <- ggplot(df, aes(x = ps, fill = hasfertilityratedatalabel, color = hasfertilityratedatalabel)) +
  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
  facet_grid(as.factor(hasfertilityratedata) ~ .) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  scale_color_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical')
print(plot2)

plot3 <- df %>%
  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
  group_by(hasfertilityratedata, ps.grp) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(n2 = ifelse(hasfertilityratedata == 0, yes = n, no =  -1*n)) %>%
  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasfertilityratedata))) +
  geom_bar(stat = 'identity', position = 'identity') +
  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasfertilityratedata == 0, 8, -8))) +
  xlab('Probability of Being in the Analysis') +
  ylab('N') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete(name = "", labels = c("Counties in States With Data", "Counties in States Without Data")) +
  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
  theme(legend.position = 'bottom', legend.direction = 'vertical',
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
print(plot3)

#run model
countydemfertilityrateanalysis <- df %>% filter(hasfertilityratedata == 1)

model <- lm(Fertility.Rate ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemfertilityrateanalysis, weights = ps)
summary(model)
